Estimating the contribution of selection to temporal allele frequency change

### 2019-09-09 manuscript/manuscript.tex
version sent to committee

\documentclass[11pt]{article}
\RequirePackage{fullpage}
\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{mathtools}
\RequirePackage{graphicx}
\RequirePackage[normalem]{ulem}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{subcaption}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{tikz}

@@

\title{Estimating the contribution of selection to temporal allele frequency change. }

@@

\begin{abstract}
%Natural selection  as evident from numerous studies demonstrating
%rapid phenotypic change
  %As we developed
%  in another paper \parencite{Buffalo2019-io}, o
  \gc{Adaptation} can proceed in natural populations and \gc{selection
  experiments over} very short timescales.
  However, detecting these changes genome-wide is difficult, as rapid
  adaptation often proceeds from standing variation or is the result of
  polygenic selection, both of which \gc{may} leave faint genomic signals
  indistinguishable from a noisy background of genetic drift. One promising signal comes from
  the \gc{genome-wide} covariance between allele frequency changes observable from temporal
  genomic data, e.g. evolve-and-resequence studies. These temporal covariances
  reflect how the change in neutral allele frequency at one timepoint is
  predictive of the changes at later timepoints when there is heritable fitness
  variation in the population, as neutral alleles \gc{can remain
  associated with their random selected background over time.}
  %become stochastically linked
 % with selected loci.
  Since genetic drift does not lead to temporal covariance,
  we can \gc{use these covariances to} begin to directly estimate what fraction of the variation in allele
  frequency change through time is drift by linked selection. Here, we
  reanalyze two \emph{Drosophila simulans} evolve-and-resequence studies, and
  one artificial selection experiment in mice, to quantify the effects of
  linked selection over short timescales. Additionally, taking advantage of
  selection on replicated populations, we can directly estimate the covariance
  in allele frequency between replicates due to convergent selection
  pressure.
  \gc{We also identify signals of negative temporal covariance that
   show that selection reverses its direction for a reasonable
   proportion of loci over the time course of a selection experiment.}
  Overall, we find that in the three studies we analyzed, linked selection has
  a remarkable impact on short term allele frequency dynamics that previously
  may have been attributed to genetic drift
\end{abstract}

@@

A long-standing problem in evolutionary genetics is quantifying the roles
genetic drift and selection have in shaping genome-wide allele frequency
changes. Selection can both directly and indirectly affect allele frequency
change, with the indirect effect coming from the action of selection on
correlated loci elsewhere in genome e.g. linked selection
(\cite{Maynard_Smith1974-lc}, \cite{Charlesworth1993-gb,Nordborg1996-nq}; see
\cite{Barton2000-zg} for a review).  Previous work on this question has mostly
focused on teasing apart the impacts of drift and selection on genome-wide
diversity using population samples from a single contemporary timepoint, often
by modeling the correlation between regional recombination rate, gene density,
and diversity created in the presence of linked selection
\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach \vb{has shown linked
  selection has a major role in shaping patterns of genome-wide diversity
across the genomes of some of organisms}
\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and \vb{negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}}.
However, we lack an understanding of how genome-wide linked selection acts over
an important dimension: time.

Presently, there are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection has on genome-wide variation
over short timescales in sexual populations remarkably difficult. This is
because the effect of selection on a polygenic trait (such as fitness) is
distributed across loci proportional to their effect sizes, which can lead to
subtle allele frequency shifts \vb{on standing variation} that are difficult to
distinguish from background levels of genetic drift and sampling variance.
However, increasingly genomic experimental evolution studies with multiple
timepoints, and in some cases multiple replicate populations, are being used to
detect selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate
modes of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.
In addition these temporal-genomic studies have begun in wild populations,
\gc{some} with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In
a previous paper, we proposed that one useful signal for understanding the
genome-wide impact of polygenic linked selection detectable from temporal
genomic studies} is the temporal autocovariance in allele frequency changes
\parencite{Buffalo2019-io}.  These covariances are directly estimable from
temporal genomic data and are created when the loci that underly heritable
fitness variation perturb linked neutral alleles; \vb{in contrast}, when
genetic drift acts alone in a closed population, these covariances are zero in
expectation. Mathematically, temporal covariances \gc{are useful} because it is
  natural to decompose the total variance in allele frequency change across a
  set of time intervals into the variances and covariances in allele frequency
  change among time intervals.  Furthermore, biologically, these covariances
  reflect the extent to which allele frequency changes in one generation
predict changes in another due to a shared fitness background and shared
selection pressures.

@@

\section{Results}

@@

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (\vb{in ten generation intervals, denoted} $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity (see Supplementary Material Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
the bias correction). The covariances are expected to be zero when only drift
is acting in a closed population, as only heritable variation for fitness can
create covariance between allele frequency changes \parencite{Buffalo2019-io}.
Averaging across the ten replicate temporal covariances matrices, we find
temporal covariances that are statistically significant (95\% block bootstraps
CIs do not contain zero), consistent with linked selection perturbing
genome-wide allele frequency changes over very short time periods. The
covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals, as expected when
directional selection affects linked variants' frequency trajectories until
ultimately linkage disequilibrium and the additive genetic variance for fitness
\vb{associated with neutral alleles} decays \parencite{Buffalo2019-io}. The
temporal covariances per replicate are noisier but this general pattern holds;
see Supplementary Figures \ref{suppfig:barghi-cov-panels}.
\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
time intervals are on average ten generations apart, and given the temporal
decay in covariance we see, the covariances on shorter time-scales (\vb{e.g. if
adjacent generations had been sequenced}) may well be higher yet (see
Supplementary Material Section \ref{supp:barghi-covs} for more details).

@@

While the presence of positive temporal covariances is consistent with linked
selection affecting allele frequencies over time, this measure not easily
interpretable. Additionally, we can quantify the impact of linked selection on
allele frequency change as the ratio of total covariance in allele frequency
change to the total variance in allele frequency change. Since the total
variation in allele frequency change can be \vb{partitioned} into variance and
covariance components, $\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i,
\Delta p_j) / \var(p_t - p_0)$, and the covariances are zero when drift acts
alone, this is a lower bound on how much of the variance in allele frequency
change is caused by linked selection \parencite{Buffalo2019-io}. We call this
measure $G(t)$, which is \vb{this measure of} the effect of linked selection
between the initial generation $0$ and some later generation $t$, which can be
varied to see how this quantity grows through time. As with the temporal
covariances, the study design of \textcite{Barghi2019-qy} leads our measure
$G(t)$ to be even more conservative, since the temporal covariances with each
ten-generation block between sequenced timepoints are not directly observable,
and are not included in the numerator of $G(t)$. Still, we find a remarkably
strong signal that greater than $20\%$ of total variation in allele frequency
change over 60 generations is directly the result of linked selection.

The replicate design of \textcite{Barghi2019-qy} also allows us to quantify
another covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances are: neutral alleles linked to a particular fitness background
experience are expected to have allele frequency changes in the same direction
if the selection pressures are similar. \vb{Intuitively, where temporal
  covariances reflect that neutral alleles associated with heritable fitness
  backgrounds are predictive of frequency changes between generations,
  replicate covariances reflect that heritable fitness backgrounds common to
each replicate predict (under the same selection pressures) frequency changes
between replicates.} We measure this through a statistic similar to a
correlation, which we call the convergent correlation: the ratio of average
between-replicate covariance across all pairs to the average standard deviation
across all pairs of replicates,

@@

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2} A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2} A). We find these convergent covariances decay very
quickly, from an initial value of about 0.1 (95\% block bootstrap confidence
intervals $[0.094, 0.11]$), to around 0.01 (95\% CIs [0.0087, 0.015]) within 20
generations.

@@

Finally, we observed that in the longest study  we analyzed
\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This shows that alleles that were on average going up
initially are later going down in frequency, i.e. that the average direction of
selection experienced by alleles has flipped. This must reflect either a change
in the environment or the genetic background, due to changing epistatic
relationships among alleles or  recombination breaking up selective alleles.
Such reversals of selective dynamics could be occurring at other timepoints but
the signal of a change in the direction of selection at particular loci may be
washed out when we calculate our genome-wide average temporal covariances.  To
address this limitation, we calculated the distribution of the temporal
covariances over 100kb genomic windows (Figure \ref{fig:figure-3}, pooling
across all replicates; see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance of each tile will be noisy, due to sampling and genetic drift, and
the neutral distribution of the covariance is complicated due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). \vb{To address
this}, we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the   allele frequency changes
per-genomic window. This destroys the systematic covariances created by linked
selection \vb{and} creates a sampling distribution of the covariances
spuriously created by neutral genetic drift while preserving the complex
dependencies between adjacent loci created by linkage disequilibrium (see
Supplementary Material Section \ref{supp:empirical-null} for more details on
this approach). This empirical neutral null distribution is conservative in
the sense that the variances of the covariances are wider than expected under
drift alone as they include the effect of selection on the allele frequency
change within a time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) that windowed temporal covariances between close
timepoints are skewed positive (a heavy right tail), while between more distant
timepoints these windowed temporal covariances tend to shift to become more
negative (a heavy left tail).  We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
striking pattern that the windowed covariances not only decay towards zero, but
in fact become negative through time, consistent with many regions in the
genome having had a reversed fitness effect at later timepoints.

@@

\begin{figure}[!htb]
  \centering
  \includegraphics[]{figures/figure-3.pdf}

  \caption{{\bf A}, {\bf B}: The distribution of temporal covariances
    calculated in 100kb genomic windows from the \textcite{Barghi2019-qy}
    study, plotted alongside an empirical neutral null distribution created by
    recalculating the windowed covariances on a 1,000 sign permutations of
    allele frequency changes within tiles. The histogram bin number is 88,
    chosen by cross validation (Supplementary Materials
    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A}, windowed
    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
    10$ generations and in subfigure {\bf A} the covariances are separated by $k=4
    \times 10$ generations; each $k$ is an off-diagonal from the variance
    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
    of covariance matrix in subfigures {\bf A} and {\bf B}, where the first diagonal is the
    variance, and the dark gray indicates which off-diagonal of the covariance
    matrix is plotted in the histograms). {\bf C}: The lower and upper tail
    probabilities of the observed windowed covariances, at 20\% and 80\%
    quintiles of the empirical neutral null distribution, for varying time
    between allele frequency changes (i.e. which off-diagonal $k$). The
  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
light gray dashed line indicates the 20\% tail probability expected under the
neutral null.}

    \label{fig:figure-3} 
\end{figure}

@@

%\gc{TODO GIVE conversion \% var due to linked selection. }

Second, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response of
selection of about 5 standard deviations over the course of twenty generations
\parencite{Castro2019-uk}. This study includes two independent selection lines,
Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
Consequently, this selection experiment offers a useful control to test our
between-replicate covariances: we expect to see positive between-replicate
covariance in the comparison between the two Longshanks selection lines, but
not between the two pairwise comparisons between the control line and the two
Longshanks lines. We find that this is the case (gray confidence intervals in
Figure \ref{fig:figure-2} C), with the two Longshanks comparisons to the
control line not being significantly different from zero, while the comparison
between the two Longshanks line is statistically significantly different from
zero (CIs [0.013, 0.037]).

A major finding in the Longshanks study were two major-effect loci that showed
parallel frequency shifts between the two selection lines: a region harboring
the gene Nkx3-2 known to be involved in limb development, and another region
harboring six other candidate genes. We were curious to what extent our
genome-wide covariances were being driven by these two outlier large-effect
loci, so we excluded them from the analysis. Since we do not know the extent to
which linkage disequilibrium around these large-effect loci affects neighboring
loci, we took the conservative precaution of excluding the entire chromosomes
these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
covariances. We find excluding these large effect loci has little impact on the
confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
indicating that we these across-replicate covariances are indeed driven by a
polygenic signal.

@@

@@

Since the seminal theoretic model of \textcite{Maynard_Smith1974-lc}
demonstrating that diversity is reduced at linked neutral variation as an
advantageous polymorphism sweeps to fixation, over four decades of theoretical
and empirical research has enriched our understanding of linked selection. This
work has led to the discovery of selected genes from genomic signals of
hitchhiking \parencite{Nair2003-tw,Voight2006-rn}, found genome-wide signals of
linked selection
\parencite{Aguade1989-jx,Begun1992-ey,Cutter2010-gi,Andersen2012-bj,Cutter2003-tl}
leading to the development of recurrent hitchhiking \parencite{Stephan1992-jc}
and background selection \parencite{Charlesworth1993-gb} models that attempt
explain this pattern. Since, much attention has focused on quantifying the
impact of background selection \parencite{McVicker2009-ax} and developing
theoretic machinery \parencite{Coop2012-cd} to differentiate the regional
reductions in diversity into those caused by full and partial sweeps, and those
caused by background selection \parencite{Elyashiv2016-vt}. However, as other
theoretic models of soft sweeps from standing variation emerged
\parencite{Hermisson2005-hs,Pennings2006-lj}, attention shifted towards
detecting other modes of linked selection \parencite{Pritchard2010-tk}, a more
difficult endeavor \parencite{Przeworski2005-bg} met with new coalescent models
\parencite{Berg2015-xj} and machine learning methods
\parencite{Schrider2017-yx}. An undervalued theoretic approach to understanding
the genome-wide effects of selection on standing variation, e.g. selection on
an infinitesimal polygenic trait, stems from an early quantitative genetic
model of linked selection \parencite{Robertson1961-ho} and its later
developments \parencite{Santiago1995-hx,Santiago1998-bs}. Implicit in these
models is that autocovariance between allele frequency change is created when
there is heritable fitness variation in the population, a signal that can be
readily detected from temporal genomic data \parencite{Buffalo2019-io}. Here,
we provide the first empirical evidence of linked selection acting on standing
variation over very short timescales.

>>>>

\documentclass[11pt]{article}
\RequirePackage{fullpage}
\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{mathtools}
\RequirePackage{graphicx}
\RequirePackage[normalem]{ulem}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{wrapfig}
\RequirePackage{subcaption}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{tikz}

@@

\title{Estimating the contribution of selection to temporal allele frequency change}

@@

\begin{abstract}

  Adaptation can proceed in natural populations and selection experiments over
  very short timescales.  However, detecting these changes genome-wide is
  difficult, as rapid adaptation often proceeds from standing variation or is
  the result of polygenic selection, both of which may leave faint genomic
  signals indistinguishable from a noisy background of genetic drift. One
  promising signal comes from the genome-wide covariance between allele
  frequency changes observable from temporal genomic data, e.g.
  evolve-and-resequence studies. These temporal covariances reflect how the
  change in neutral allele frequency at one timepoint is predictive of the
  changes at later timepoints when there is heritable fitness variation in the
  population, as neutral alleles can remain associated with their random 
  selected background over time. Since genetic drift does not lead to temporal
  covariance, we can use these covariances to begin to directly estimate what
  fraction of the variation in allele frequency change through time is drift by
  linked selection. Here, we reanalyze two \emph{Drosophila simulans}
  evolve-and-resequence studies, and one artificial selection experiment in
  mice, to quantify the effects of linked selection over short timescales.
  Additionally, taking advantage of selection on replicated populations, we can
  directly estimate the covariance in allele frequency between replicates due
  to convergent selection pressure.  We also identify signals of negative
  temporal covariance that show that selection reverses its direction for a
  reasonable proportion of loci over the time course of a selection experiment.
  Overall, we find that in the three studies we analyzed, linked selection has
  a remarkable impact on short term allele frequency dynamics that previously
  may have been attributed to genetic drift

\end{abstract}

@@

A long-standing problem in evolutionary genetics is quantifying the roles
genetic drift and selection have in shaping genome-wide allele frequency
changes. Selection can both directly and indirectly affect allele frequency
change, with the indirect effect coming from the action of selection on
correlated loci elsewhere in genome e.g. linked selection
(\cite{Maynard_Smith1974-lc}, \cite{Charlesworth1993-gb,Nordborg1996-nq}; see
\cite{Barton2000-zg} for a review).  Previous work on this question has mostly
focused on teasing apart the impacts of drift and selection on genome-wide
diversity using population samples from a single contemporary timepoint, often
by modeling the correlation between regional recombination rate, gene density,
and diversity created in the presence of linked selection
\parencite{Begun1992-ey,Elyashiv2016-vt}. This approach has shown linked
selection has a major role in shaping patterns of genome-wide diversity across
the genomes of some of organisms
\parencite{Begun2007-bg,Beissinger2016-cm,Sattath2011-dr,Williamson2014-oy,Andersen2012-bj,Cutter2010-gi},
and has allowed us to quantify the relative influence of positive selection
(hitchhiking) and negative selection (background selection;
\cite{Nordborg2005-dc,McVicker2009-ax,Hernandez2011-gs,Elyashiv2016-vt}.
However, we lack an understanding of how genome-wide linked selection acts over
an important dimension: time.

Presently, there are numerous examples of rapid phenotypic adaptation
\parencite{Grant2011-wk,Grant2006-hj,Reznick1997-mh,Franks2007-dr} and rapid
genomic evolution in asexual populations
\parencite{Good2017-om,Bennett1990-bc,Baym2016-kh}.  Yet the polygenic nature
of fitness makes detecting the impact selection has on genome-wide variation
over short timescales in sexual populations remarkably difficult. This is
because the effect of selection on a polygenic trait (such as fitness) is
distributed across loci proportional to their effect sizes, which can lead to
subtle allele frequency shifts on standing variation that are difficult to
distinguish from background levels of genetic drift and sampling variance.
However, increasingly genomic experimental evolution studies with multiple
timepoints, and in some cases multiple replicate populations, are being used to
detect selected loci \parencite{Turner2011-sx,Turner2012-bm} and differentiate
modes of selection \parencite{Burke2010-tz,Barghi2019-qy,Therkildsen2019-zy}.
In addition these temporal-genomic studies have begun in wild populations, some
with the goal of finding variants that exhibit frequency changes consistent
with fluctuating selection \parencite{Bergland2014-ij,Machado2018-cs}. \vb{In a
  previous paper, we proposed that one useful signal for understanding the
  genome-wide impact of polygenic linked selection detectable from temporal
genomic studies} is the temporal autocovariance in allele frequency changes
\parencite{Buffalo2019-io}.  These covariances are directly estimable from
temporal genomic data and are created when the loci that underly heritable
fitness variation perturb linked neutral alleles; in contrast, when genetic
drift acts alone in a closed population, these covariances are zero in
expectation. Mathematically, temporal covariances are useful because it is
natural to decompose the total variance in allele frequency change across a set
of time intervals into the variances and covariances in allele frequency change
among time intervals.  Furthermore, biologically, these covariances reflect the
extent to which allele frequency changes in one generation predict changes in
another due to a shared fitness background and shared selection pressures.

@@

\clearpage
\section{Results}

@@

We first analyzed \textcite{Barghi2019-qy}, an evolve-and-resequence study with
ten replicate populations exposed to a high temperature lab environment and
evolved for 60 generations, and sequenced every ten generations. Using the
seven timepoints and ten replicate populations, we estimated the genome-wide $6
\times 6$ temporal covariance matrix $\mathbf{Q}$ for each of the ten
replicates. Each row of these matrices represent the temporal covariance
$\cov(\Delta_{_{10}} p_s, \Delta_{_{10}} p_t)$, between the allele frequency
change (in ten generation intervals, denoted $\Delta_{_{10}} p_t$) in some
initial reference generation $s$ (the row of the matrix), and some later
timepoint $t$ (the column of the matrix). We corrected these matrices for
biases created due to sampling noise, and normalize the entries for
heterozygosity (see Supplementary Material Sections
\ref{supp:ind-depth-var-corr} and \ref{supp:matrix-correction} for details on
the bias correction). The covariances are expected to be zero when only drift
is acting in a closed population, as only heritable variation for fitness can
create covariance between allele frequency changes \parencite{Buffalo2019-io}.
Averaging across the ten replicate temporal covariances matrices, we find
temporal covariances that are statistically significant (95\% block bootstraps
CIs do not contain zero), consistent with linked selection perturbing
genome-wide allele frequency changes over very short time periods. The
covariances between all adjacent time intervals are positive and then decay
towards zero as we look at more distant time intervals, as expected when
directional selection affects linked variants' frequency trajectories until
ultimately linkage disequilibrium and the additive genetic variance for fitness
associated with neutral alleles decays \parencite{Buffalo2019-io}. The temporal
covariances per replicate are noisier but this general pattern holds; see
Supplementary Figures \ref{suppfig:barghi-cov-panels}.
\textcite{Barghi2019-qy}'s design means that the covariances we see in adjacent
time intervals are on average ten generations apart, and given the temporal
decay in covariance we see, the covariances on shorter time-scales (e.g. if
adjacent generations had been sequenced) may well be higher yet (see
Supplementary Material Section \ref{supp:barghi-covs} for more details).

@@

While the presence of positive temporal covariances is consistent with linked
selection affecting allele frequencies over time, this measure not easily
interpretable. Additionally, we can quantify the impact of linked selection on
allele frequency change as the ratio of total covariance in allele frequency
change to the total variance in allele frequency change. Since the total
variation in allele frequency change can be partitioned into variance and
covariance components, $\var(p_t - p_0) = \sum_{i \ne j} \cov(\Delta p_i,
\Delta p_j) / \var(p_t - p_0)$, and the covariances are zero when drift acts
alone, this is a lower bound on how much of the variance in allele frequency
change is caused by linked selection \parencite{Buffalo2019-io}. We call this
measure $G(t)$, which is this measure of the effect of linked selection
between the initial generation $0$ and some later generation $t$, which can be
varied to see how this quantity grows through time. As with the temporal
covariances, the study design of \textcite{Barghi2019-qy} leads our measure
$G(t)$ to be even more conservative, since the temporal covariances with each
ten-generation block between sequenced timepoints are not directly observable,
and are not included in the numerator of $G(t)$. Still, we find a remarkably
strong signal that greater than $20\%$ of total variation in allele frequency
change over 60 generations is directly the result of linked selection.

The replicate design of \textcite{Barghi2019-qy} also allows us to quantify
another covariance: the covariance in allele frequency change between replicate
populations experiencing convergent selection pressures. These
between-replicate covariances are created in the same way as temporal
covariances: neutral alleles linked to a particular fitness background
experience are expected to have allele frequency changes in the same direction
if the selection pressures are similar. Intuitively, where temporal covariances
reflect that neutral alleles associated with heritable fitness backgrounds are
predictive of frequency changes between generations, replicate covariances
reflect that heritable fitness backgrounds common to each replicate predict
(under the same selection pressures) frequency changes between replicates. We
measure this through a statistic similar to a correlation, which we call the
convergent correlation: the ratio of average between-replicate covariance
across all pairs to the average standard deviation across all pairs of
replicates,

@@

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2} A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2} A). We find these convergent covariances are
relatively weak, and decay very quickly from an initial value of about 0.1
(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
(95\% CIs [0.0087, 0.015]) within 20 generations. This seems to confirm a
primary finding of the original \textcite{Barghi2019-qy} study: that
alternative loci contribute to adaptation in different replicates.

@@

@@

@@

%\gc{TODO GIVE conversion \% var due to linked selection. }
Second, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response of
selection of about 5 standard deviations over the course of twenty generations
\parencite{Castro2019-uk}. This study includes two independent selection lines,
Longshanks 1 and 2 (LS1 and LS2), and an unselected control line (Ctrl).
Consequently, this selection experiment offers a useful control to test our
between-replicate covariances: we expect to see positive between-replicate
covariance in the comparison between the two Longshanks selection lines, but
not between the two pairwise comparisons between the control line and the two
Longshanks lines. We find that this is the case (gray confidence intervals in
Figure \ref{fig:figure-2} C), with the two Longshanks comparisons to the
control line not being significantly different from zero, while the comparison
between the two Longshanks line is statistically significantly different from
zero (CIs [0.013, 0.037]).

A major finding in the Longshanks study were two major-effect loci that showed
parallel frequency shifts between the two selection lines: a region harboring
the gene Nkx3-2 known to be involved in limb development, and another region
harboring six other candidate genes. We were curious to what extent our
genome-wide covariances were being driven by these two outlier large-effect
loci, so we excluded them from the analysis. Since we do not know the extent to
which linkage disequilibrium around these large-effect loci affects neighboring
loci, we took the conservative precaution of excluding the entire chromosomes
these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
covariances. We find excluding these large effect loci has little impact on the
confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
indicating that these across-replicate covariances are indeed driven by a
polygenic signal.

@@

\begin{figure}
  \begin{center}
  \includegraphics[]{figures/figure-3.pdf}

  \caption{\footnotesize {\bf A}, {\bf B}: The distribution of temporal covariances
    calculated in 100kb genomic windows from the \textcite{Barghi2019-qy}
    study, plotted alongside an empirical neutral null distribution created by
    recalculating the windowed covariances on a 1,000 sign permutations of
    allele frequency changes within tiles. The histogram bin number is 88,
    chosen by cross validation (Supplementary Materials
    \ref{suppfig:barghi-cross-validation-binsize}). In subfigure {\bf A}, windowed
    covariances $\cov(\Delta p_t, \Delta p_{t+k})$ are separated by $k=2 \times
    10$ generations and in subfigure {\bf A} the covariances are separated by $k=4
    \times 10$ generations; each $k$ is an off-diagonal from the variance
    diagonal of the temporal covariance matrix (see cartoon of upper-triangle
    of covariance matrix in subfigures {\bf A} and {\bf B}, where the first diagonal is the
    variance, and the dark gray indicates which off-diagonal of the covariance
    matrix is plotted in the histograms). {\bf C}: The lower and upper tail
    probabilities of the observed windowed covariances, at 20\% and 80\%
    quintiles of the empirical neutral null distribution, for varying time
    between allele frequency changes (i.e. which off-diagonal $k$). The
  confidence intervals are  $95\%$ block-bootstrap confidence interval, and the
light gray dashed line indicates the 20\% tail probability expected under the
neutral null.}
    \end{center}
    \label{fig:figure-3} 
\end{figure}

\clearpage

Finally, we observed that in the longest study  we analyzed
\parencite{Barghi2019-qy}, some genome-wide temporal covariances become
negative at future timepoints (see the first two rows in Figure
\ref{fig:figure-1} A). This shows that alleles that were on average going up
initially are later going down in frequency, i.e. that the average direction of
selection experienced by alleles has flipped. This must reflect either a change
in the environment or the genetic background, due to changing epistatic
relationships among alleles or  recombination breaking up selective alleles.
Such reversals of selective dynamics could be occurring at other timepoints but
the signal of a change in the direction of selection at particular loci may be
washed out when we calculate our genome-wide average temporal covariances.  To
address this limitation, we calculated the distribution of the temporal
covariances over 100kb genomic windows (Figure \ref{fig:figure-3}, pooling
across all replicates; see Supplementary Figure
\ref{suppfig:barghi-offset-replicate-panels} for individuals replicates). The
covariance of each tile will be noisy, due to sampling and genetic drift, and
the neutral distribution of the covariance is complicated due to linkage
disequilibria (which can occur over long physical distances in E\&R and
selection studies, \cite{Nuzhdin2013-gf,Baldwin-Brown2014-cl}). To address
this, we have developed a permutation-based procedure that constructs a null
distribution by randomly flipping the signs of the  allele frequency changes
per-genomic window. This destroys the systematic covariances created by linked
selection and creates a sampling distribution of the covariances
spuriously created by neutral genetic drift while preserving the complex
dependencies between adjacent loci created by linkage disequilibrium (see
Supplementary Material Section \ref{supp:empirical-null} for more details on
this approach). This empirical neutral null distribution is conservative in
the sense that the variances of the covariances are wider than expected under
drift alone as they include the effect of selection on the allele frequency
change within a time-interval, just not between time-intervals. We see (Figure
\ref{fig:figure-3} A and B) that windowed temporal covariances between close
timepoints are skewed positive (a heavy right tail), while between more distant
timepoints these windowed temporal covariances tend to shift to become more
negative (a heavy left tail).  We quantified the degree to which the left and
right tails are inflated compared to the null distribution as a function of
time, and see excesses in both tails in Figure \ref{fig:figure-3} C. This
finding is also robust to sign-permuting allele frequency changes on a
chromosome-level, the longest extent that gametic linkage disequilibria can
extend (Supplementary Figure \ref{suppfig:barghi-tailprobs-seqid}). We see a
striking pattern that the windowed covariances not only decay towards zero, but
in fact become negative through time, consistent with many regions in the
genome having had a reversed fitness effect at later timepoints.

@@

Since the seminal theoretic model of \textcite{Maynard_Smith1974-lc}
demonstrating that diversity is reduced at linked neutral variation as an
advantageous polymorphism sweeps to fixation, over four decades of theoretical
and empirical research has enriched our understanding of linked selection. This
work has led to the discovery of selected genes from genomic signals of
hitchhiking \parencite{Nair2003-tw,Voight2006-rn}, found genome-wide signals of
linked selection
\parencite{Aguade1989-jx,Begun1992-ey,Cutter2010-gi,Andersen2012-bj,Cutter2003-tl}
leading to the development of recurrent hitchhiking \parencite{Stephan1992-jc}
and background selection \parencite{Charlesworth1993-gb} models that attempt
explain this pattern. Since, much attention has focused on quantifying the
impact of background selection \parencite{McVicker2009-ax} and developing
theoretic machinery \parencite{Coop2012-cd} to differentiate the regional
reductions in diversity into those caused by full and partial sweeps, and those
caused by background selection \parencite{Elyashiv2016-vt}. However, as other
theoretic models of soft sweeps from standing variation emerged
\parencite{Hermisson2005-hs,Pennings2006-lj}, attention shifted towards
detecting other modes of linked selection \parencite{Pritchard2010-tk}, a more
difficult endeavor \parencite{Przeworski2005-bg} met with new coalescent models
\parencite{Berg2015-xj} and machine learning methods
\parencite{Schrider2017-yx}. An undervalued theoretic approach to understanding
the genome-wide effects of selection on standing variation, e.g. selection on
an infinitesimal polygenic trait, stems from an early quantitative genetic
model of linked selection \parencite{Robertson1961-ho} and its later
developments (\cite{Santiago1995-hx,Santiago1998-bs}; see also
\cite{Barton2000-zg} for a comparison of these models with classic hitchhiking
models). Implicit in these models is that autocovariance between allele
frequency change is created when there is heritable fitness variation in the
population, a signal that can be readily detected from temporal genomic data
\parencite{Buffalo2019-io}. Here, we provide the first empirical evidence of
linked selection acting on standing variation over very short timescales.

We reanalyzed three published temporal genomic datasets and detected the
genome-wide signature of linked selection acting over tens of generations.
Furthermore, we find the dynamics of temporal autocovariance are consistent
with theory; the temporal covariance in allele frequencies decays as both
the fitness variation associated with particular haplotypes and the linkage
disequilibria between selected and neutral loci decay through time. In
reanalyzing one study \parencite{Barghi2019-qy}, we find that after sixty
generations, greater than 20\% of the variation in allele frequency change is
directly due to the action linked selection. Capitalizing on replicated
evolve-and-resequence study designs, we characterized the extent to which
convergent selection pressures lead to parallel changes in allele frequency
across replicate populations, finding moderate correlations in allele frequency
changes.
